Vragenlijst thesis evaluatie

### 2015-06-11 doc/template/chap-10.tex
Kleine verbetering + virus prop model
Aanvullingen overal.

\subsubsection{Complication}
Since it is so difficult to protect a system or a network against APT's, researchers have been looking for effective ways to predict in advance which defence strategy might be the better one. 
Game theory is gaining increasing interest as an effective technique to model and study cyber security problems. It is common to model cyber security games as a game with two players, an attacker and a defender. However, there are games that have more players \cite{fengstealthy}. In this paper we focus on the game with two players. The actions available to the attacker and the defender correspond to attacks on the system and defensive measures that protect the system. In particular, both players will execute the strategy that results in a maximal benefit for them.  Researchers at RSA, van Dijk et al,  made a game theoretic model of targeted attacks. They study the specific scenario where a system or network is repeatedly taken over completely by an without being immediately detected by the defender of the system or network. In game theory, such a game is known as ''FlipIt'' \cite{FlipIt}. This is a two players game where the attacker and the defender are competing to get control over a shared resource. Neither player knows who is currently in control of the resource until they move. In FlipIt every move gives them immediate control over the resource. But what if the attacker moves and it takes a while before the attacker gets full control over the resource? FlipIt does not take into account that a move may not be instantaneous, but has a certain delay. Consider for example a network with different nodes ( laptops, datacenters) as a resource. The attacker drops a virus on one of the nodes and then wait till this virus infects the whole network. The attacker will only be in control of the resource once the whole network is infected. \\

@@

\subsubsection{Contributions}
We propose an addition to the basic FlipIt model to model a scenario where the moves by the attacker will not be instantaneous. Next we analyse what the new Nash equilibria will be and .. \\

\subsubsection{Overview of the thesis}

>>>>

\subsubsection{Complication}
Since it is so difficult to protect a system or a network against APT's, researchers have been looking for effective ways to predict in advance which defence strategy might be the better one. 
Game theory is gaining increasing interest as an effective technique to model and study cyber security problems. It is common to model cyber security games as a game with two players, an attacker and a defender. However, there are games that have more players \cite{fengstealthy}. In this paper we focus on the game with two players. The actions available to the attacker and the defender correspond to attacks on the system and defensive measures that protect the system. In particular, both players will execute the strategy that results in a maximal benefit for them.  Researchers at RSA, van Dijk et al,  made a game theoretic model of targeted attacks. They study the specific scenario where a system or network is repeatedly taken over completely by an without being immediately detected by the defender of the system or network. In game theory, such a game is known as ``FlipIt'' \cite{FlipIt}. This is a two players game where the attacker and the defender are competing to get control over a shared resource. Neither player knows who is currently in control of the resource until they move. In FlipIt every move gives them immediate control over the resource. But what if the attacker moves and it takes a while before the attacker gets full control over the resource? FlipIt does not take into account that a move may not be instantaneous, but has a certain delay. Consider for example a network with different nodes (laptops, datacenters) as a resource. The attacker drops a virus on one of the nodes and then wait till this virus infects the whole network. The attacker will only be in control of the resource once the whole network is infected. \\

@@

\subsubsection{Contributions}
The following contributions are made in this paper:
\begin{itemize}
\item[-] We propose an addition to the basic FlipIt model to model a scenario where the moves by the attacker will not be instantaneous. We extend the FlipIt game where the attacker flips with a delay. The attacker only comprimises the system if suffi"ent nodes in the network are infected. 
\item[-] Another derivation of the method to calculate the benefit of the basic FlipIt game is given.
\item[-] The periodic case of FlipIt is modelled with a delay, resulting in optimum functions and nash equilibria. 
\item[-] Based on our results we can give practical recommendations to take measures against advanced attacks. 
\end{itemize}
 Next we analyse what the new Nash equilibria will be and .. \\

While it seems maybe trivial to extend FlipIt with a delay it is not so straightforward. First a formula to calculate the benefit of a basic FlipIt was derived to introduce the delay. To calculate the benefit for each case \todo{cases misschien uitleggen?}, the intervals where not independant anymore as in the basic FlipIt game. Er moest rekening gehouden worden met de vorige intervallen om de benefit te berekenen in het eigen interval. 
\subsubsection{Overview of the thesis}

### 2015-06-11 doc/template/chap-4.tex
Aanvullingen overal.

%\chapter{Intoduction to GameTheory}
%\label{cha:1}
\documentclass[10pt]{article}
\begin{document}

\section{Write down the settings of the game}

@@

\end{document}

>>>>

\chapter{Virus Propagation }
\label{chapter4: Virus propagation}

\section{Methods of virus propagation}

The spreading of worms has already been extensivly researched [beter schrijven]. Because of all the different types of propagation it is hard to pick out one model that can model all of them. Modelling the spread of malware depends on two great factors: the method used for the propagation and the graph where the malware will spread itself. Viruses and worms are the types of malware that are the most researched. Under we list a couple of examples to model virus and worm propagation. In this text worm and virus will be used intermediate.

\subsection{Models of graph representing the network}

\subsection*{Methods of propagation}

\begin{description}
\item Random scanning: Some worms propagate by using the method of random scanning 
IP addresses.  Rate of terms of succeeded randwomly choosen IP adresses -> maximise good ip adresses. After a while the remaining nodes will be les likely to be reached. An example of such a worm is `Code Red' .
\item Routing worms: they use BGP routing tables to scan the routable address space. 3 times faster than traditional worm that uses random scanning. Spyb0t or network bluepill.
\item email worms: use the email systems to propagate. The Kak worm is a Javascript computer worm that spread itself by exploiting a bug in Outlook Express. 
%Pikachu worm: The virus was mainly spread through Microsoft Outlook email attachments. The email containing the attached virus propagated through infected users by sending itself to all contacts in the user's Outlook address book. [5]
\item self stopping worm: reduces speed to avoid detection. Atak worm or self stopping worm
\end{description}

The start of such a propagation can be different. Either it is a targeted attack or a random attack. For example a usb stick or other shared resources can be used to start the spreading. It can be given to some special person or it can just be left behind in a parking lot. This can still be a targeted attack if it is the parking lot of a targeted company but there is more insecurity of where the usb stick will end. Another way is by sending an email. It can also start from one computer that is connected directly to the internet. Downloading malicious programs.

\subsection*{Models for worm propagation}
To use the models for worm propagation in our model of the FlipIt game it is suffi"ent to know the propagation speed of the worm and define after how many time units the network is defected. This will be equal to the \textit{d} parameter in our model.  
Stackelberg game: first move is from attacker, defender is follower.

Kind of network model: 
Kind of model: SIS, SIR

\subsubsection*{SIS}

\subsubsection*{SIR}

\subsubsection*{self disciplinary worms}
Model for self disciplinary worms and counter measures ... []

Popular mechanism that worms use to detect vulnerable targets by random ip scanning probing. Feasible due to use 32-bit addresses. 128-bit adresses life harder for worms, except the ones that use email systems to propagate. two new strategies: uniformly distributed random number generator to select new target. :spread locally, by biasing the search space towards addresses within the same subnet or network. 
The second strategy is almost the same as what email virusses would get. For this reason we work an example out ..

A method to calculate the propagation of the virus in an easy way. Google page ranking algorithm.

\section{methode met matrixen}

@@

What do we need for an algorithm
\begin{description}
\item Graph network $G = < V, E>$
\item Graph matrix $[A]$ which is $|V| \times |V| $
\item Attack vector $[X]$ which is $1 \times |V|$
\item cummulative matrix $[M]$ which is $|V| \times |V|$
\item state matrix $[T]$  which is $|V| \times |V|$
\item Reset vector $[R]$
\item duration \textit{d}
\item time \textit{n}
\item rate $\delta _{0}$ of defender and $\delta _{1}$ of attacker
\end{description}

Initialisation algorithm:

\begin{verbatim}
initialisatie
	d=0
	A=basismatrix
	M=A^{0}
	n=0
	\delta_{0}
	\delta_{1}
	X
	R
	controller = defender

	Algorithm
	n:= n + 1;
	Check who is in control? ( through modulo )
	if ( defender & controller=defender)
				d:= d + 1;

	if ( defender & controller=attacker )
				G = X \times R  (flippen ten voordele van defender)
				d = 0
				controller = defender

	if ( attacker & controller=defender )
				controller=attacker
				..

	if ( attacker & contoller=attacker )
				d:= d + 1
				M = M x A
				T = T + M
				G = X x T

\end{verbatim}

%\end{document}

### 2015-06-11 doc/template/chap-5.tex
Aanvullingen overal.

\chapter{Intoduction to GameTheory}
\label{Chapter1:Intro.Game.Theory}
%\documentclass[10pt]{article}
%\begin{document}

@@

The generic term for data protection stored on a computer controlled device such as computers and smartphones, as well as public and private computer networks, including the entire Internet is called computer security. 
Security is a general term that encompasses several dimensions. More specifically, computer security has three key objectives that are fundamental to computer security:

@@

 %The purpose of security is to give certainty that data will not be removed without authorization (Confidentiality) that the data is always accessible (Availability), and that data can not be read or altered by someone who does not have the authorization (Integrity). 
 ~~\\
 These key attributes are also known as the CIA triad. They are the fundamental security objectives for securing data, information and computing services. \\
 \\
A subset of computer security is cyber security. This is the process of applying security tools to ensure confidentiality, integrity, and availability of data.\todo{what is cyber security? In deze tekst lijkt het op computer security}

@@

@@

According to a security survey of Kaspersky \cite{SurveyKaspersky}  the damage of one successful targeted attack against a large company can exceed over 2.54 million dollar. Companies need a defence mechanism to defend themselves against APT's. As previously stated, simple detection and identification tools are insufficient to protect oneself against APT's, and a removal tool will only work if the threat has been identified. To mitigate these kind of attacks another security countermeasure is needed. This is where the researchers of RSA come into the picture. They have modelled these targeted attacks using Game Theory, and - given a certain amount of information - determined optimal defence strategies.\todo{bekijk dit}

>>>>

\chapter{Intoduction to the general context of the work}
\label{Chapter1:Intro.Game.Theory}
%\documentclass[10pt]{article}
%\begin{document}

@@

The generic term for data protection stored on a computer controlled device such as computers and smartphones, as well as public and private computer networks, including the entire Internet is called information security. 
Security is a general term that encompasses several dimensions. More specifically, information security has three key objectives that are fundamental to information security:

@@

 %The purpose of security is to give certainty that data will not be removed without authorization (Confidentiality) that the data is always accessible (Availability), and that data can not be read or altered by someone who does not have the authorization (Integrity). 
 ~~\\
 These key attributes are also known as the CIA triad. They are the fundamental security objectives for securing data, information and computing services. \\
 \\
Information security is divided into two main areas of security. One of them is cyber security also known as computer security. This is the process of applying security tools to ensure confidentiality, integrity, and availability of data. the other area is Network or Internet security.. The difference between these two terms has been lost over time. \todo{what is cyber security? In deze tekst lijkt het op computer security}

@@

%\begin{figure}[hbtp]
%\caption{2014 survey results by Kaspersky}
%\centering
%\includegraphics[scale=0.5]{Images/kaspmalware.png}
%\end{figure}

@@

According to a security survey of Kaspersky \cite{SurveyKaspersky}  the damage of one successful targeted attack against a large company can exceed over 2.54 million dollar. Companies need a defence mechanism to defend themselves against APT's. As previously stated, simple detection and identification tools are insufficient to protect oneself against APT's, and a removal tool will only work if the threat has been identified. To mitigate these kind of attacks another security countermeasures are needed. This is where the researchers of RSA come into the picture. They have modelled the situation where targeted attacks are taking over a system and are not immediately detected by the defender.

### 2015-06-11 doc/template/chap-9.tex
Aanvullingen overal.

Intuitively we could assume that $d$ can never be bigger than $\delta_{A}$ because then the benefit for the defender would always be 1. This is not always true. It is only true if $d$ is bigger than $\delta_{D}$, because then the defender will always be in control. For this we only calculate the formulas for the cases where $d$ is smaller than $\delta_{D}$. We can already conclude that it is no use for the attacker to play when the delay is bigger than $\delta_{D}$. \\

@@

@@

\subsubsection*{\textbf{Case a:} $d + \delta_{A} \leq \delta_{D}$}

The total gain  rate of the defender is then the probability that the defender will move during an interval of the attacker multiplied by the total average gain per interval:

>>>>

@@

Intuitively we could assume that $d$ can never be bigger than $\delta_{A}$ because then the benefit for the defender would always be 1. This is not always true. It is only true if $d$ is bigger than $\delta_{D}$, because then the defender will always be in control. For this we only calculate the formulas for the cases where $d$ is smaller than $\delta_{D}$. We can already conclude that it is no use for the attacker to play when the delay is bigger than $\delta_{D}$. see figure \ref{delaye}\\

\todo{figuur juist maken}
\begin{figure}[hbtp]
\caption{Delay bigger than deltaA and delay bigger than deltaD}
\centering
\includegraphics[scale=0.7]{Images/DiffDelayCase1.png}
\label{delaye}
\end{figure}

@@

\subsubsection*{\textbf{Case a:} $d + \delta_{A} \leq \delta_{D}$}
The first case is where the delay is shorter than the difference between the attacker's and the defender's period so their is no problem of counting the delay twice. To determine the total gain rate of the defender we need to know what the probability is that the defender will move during an interval and what the average time is that the defender controls the resource. \todo{beter verwoorden}.
The total gain  rate of the defender is then the probability that the defender will move during an interval of the attacker multiplied by the total average gain per interval:
